---
title: "spotifywRapped: Unwrapped!"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spotifywRapped: Unwrapped!}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The spotifywRapped package is a great tool for visualizing a user's Spotify listening history. The package comes equipped with several preset themes, called "vibes", that a user can choose to customize their visualization experience. Moreover, the listener can choose whether they want to view their long-term, medium-term, or short-term listening history. There are several kinds of built-in visualization options a user has at their disposal--let's give them a look!

## Energy & Valence Plots

The energy vs valence plot gives a visual summary of the general "vibe" a user tends towards in their favorite music. Using the user's saved tracks, the plot shows the distribution of songs across four quadrants, each representing a different mood.

The function takes the user's listening data, vibe choice, filename, and save location as arguments. It outputs a .png file to the save location of the user's choice and returns the image's file path.

```{r enval}
library(spotifywRapped)

temp_save <- tempdir()
energy_valence_demo <- my_energy_valence(saveto = temp_save, vibe = "neon", name = "enval_neon")

```

Here is an example of the energy vs valence plot:

![](`r energy_valence_demo`)

```{r, echo = FALSE}
```

```{r, echo = FALSE}

```
