---
title: "spotifywRapped: Unwrapped"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spotifywRapped: Unwrapped}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Github Repository

You can access `spotifywRapped`'s Github repository by following this link: https://github.com/Jaredhileman/spotifywRapped

## Overview

The spotifywRapped package is a valuable tool for visualizing a user's Spotify listening habits and history. The package comes equipped with several preset themes, called "vibes", that a user can choose to customize their visualization experience. These themes were created as .pngs in Canva and placed in the `inst/vibe` folder within the internal package structure. The listener can choose whether they want to view their long-term, medium-term, or short-term listening history or information about their saved (liked) tracks. There are several visualization and graphic generation functions available for the user--let's take a look at how to use them.

## Data Processing

In order to use spotifywRapped, the user must get their Spotify data by interfacing with the Spotify API. Here are the steps a user must take to organize their data for use by `spotifywRapped`.

1. The user must have Spotify developer account. Follow the instructions for creating a dev account linked here:
https://developer.spotify.com/documentation/web-api. Save the client ID and client secret for use in Step 2.
  - When setting a URI, we recommend using "http://localhost:1410/".
2. Use the `spotifyr` package (published to CRAN) to generate access and authorization tokens. The user may choose to set environment variables to include these tokens (avoids calling tokens in each function), as shown below. Otherwise, the user must include the tokens in each function call. Scopes must be specified while getting the authorization token.  `spotifywRapped` requires the "user-top-read" and "user-library-read" access. Documentation for `spotifyr` can be found here: https://cran.r-project.org/web/packages/spotifyr/spotifyr.pdf

```{r install_spotifyr, eval=FALSE}
install.packages("spotifyr")
library(spotifyr)

Sys.setenv(SPOTIFY_CLIENT_ID = "your_client_id") # From dev dashboard
Sys.setenv(SPOTIFY_CLIENT_SECRET = "your_client_secret") # From dev dashboard

access_token <- get_spotify_access_token()
authorization_code <- get_spotify_authorization_code(
  scope = c(
    "user-top-read",
    "user-library-read"
  )
)
#
```
3. The user is able to use the `get_top_artists_or_tracks()` and `get_my_saved_tracks()` functions from `spotifyr` to get their listening data. The user can then save this data to an .rda file for use in the `spotifywRapped` package. Note that both these functions have a parameter called `offset`. Since the Spotify API only allows for a limit of 50 tracks / artists per call, the user must use the offset term to get all of the saved and top tracks from the data. Please see below for a simple approach to pull the data:

```{r offset_example, eval=FALSE}
library(dplyr)

# offset value
offset_length <- ceiling(spotifyr::get_my_saved_tracks(
  include_meta_info = TRUE,
  authorization = authorization
)[["total"]] / 50)
offset_total <- (seq(offset_length) - 1) * 50

# Obtain saved tracks
saved_tracks <- data.frame()
for (batch in offset_total) {
  curr_saved_tracks <- spotifyr::get_my_saved_tracks(
    limit = 50,
    offset = batch, authorization = authorization
  )
  saved_tracks <- dplyr::bind_rows(saved_tracks, curr_saved_tracks)
}
```


4. Neither `energy` or `valence` are included in the saved_tracks data. The user must use the `get_track_audio_features()` function from the `spotifyr` package with each track's unique Spotify ID to get this information. The user can then merge this data with their saved tracks data to get a complete dataset. To get around the Spotify API rate limits, the user can use a batch size of 100 tracks per call and a sleep window of 30 seconds (corresponds to the Spotify API sliding window). Here is an example (assumes the user already have saved tracks in a datafame):

```{r get_energy_valence, eval=FALSE}
st <- saved_tracks

for (i in seq_len(st)) {
  if (i %% 100 == 0) {
    print(paste0("Sleeping... ", i %/% 100))
    Sys.sleep(30)
  }
  track_id <- st$track.id[i]
  track_features <- spotifyr::get_track_audio_features(track_id,
    authorization = access_token
  )
  st$energy[i] <- track_features$energy
  st$valence[i] <- track_features$valence
}
```

5. The user's Spotify API data should be appropriately processed and is ready for use in `spotifywRapped`.

NOTE: The `spotifyr` package nicely compacts the requested data into a dataframe. For insight into the specific structure of these datasets, please refer to the documentation for `spotifywRapped::saved_tracks` and `spotifywRapped::top_artists_longterm`.

## Energy & Valence Plots

The energy vs valence plot gives a visual summary of the general "vibe" a user tends to enjoy in their music. Using the user's saved tracks, the plot shows the distribution of songs across four quadrants, each representing a different mood.  Energy and valence are measured between 0 and 1, so values below 0.5 are considered "low valence" and "low energy", while values above 0.5 are considered "high valence" and "high energy".  The scatter plot of energy vs. valence is printed, as well as the mood name of the quadrant with the most data points.

The function takes the user's listening data, vibe choice, filename, and save location as arguments. It outputs a .png file to the save location of the user's choice and returns the image's file path.

Parameter options:

- `data`: the user's saved track data. If none specified, defaults to `spotifywRapped::saved_tracks`.

- `vibe`: "soft", "neon", "neutral", or "bright"

- `name`: the name of the output file

- `saveto`: the directory to save the output file

Here is an example of the energy vs valence plot:

```{r enval, echo=TRUE}
library(spotifywRapped)

temp_save <- tempdir()
energy_valence_demo <- my_energy_valence(data = spotifywRapped::saved_tracks, 
                                         saveto = temp_save, vibe = "bright", 
                                         name = "enval_neon")
```
![](`r energy_valence_demo`){width=450px height=800px}

## Top Artists & Tracks Graphics

Another interesting visual implemented in the spotifywRapped package is the `my_top_five` function graphic. This visualization shows the user's top five most frequently listened to artists and tracks over all time, the past 6 months, or the past 4 weeks. It prints the name of each of the top five tracks or artists and an image of the #1 song/artist.

Parameter options:

- `time`: "long", "medium", or "short"

- `vibe`: "soft", "neon", "neutral", or "bright"

- `category`: "artists" or "songs"

- `name`: the name of the output file

- `saveto`: the directory to save the output file

- `data`: the user's listening data. If none specified, defaults to package data based on `time` and `category` arguments.


The function will return the file path to the generated .png file.


Here is an example of the top artist graphic from long-term listening history:

```{r top_artists}
my_top_artists <- my_top_five(
  time = "long", vibe = "neutral", category = "artists",
  name = "My_Top_Artists_longterm",
  saveto = temp_save
)
```

![](`r my_top_artists`){width=450px height=800px}

Similarly, here is an example of the top tracks graphic from short-term listening:

```{r top_tracks}
my_top_tracks <- my_top_five(time = "short", vibe = "neon", category = "songs",
                             name = "my_top_tracks_shortterm", 
                             saveto = temp_save)
```
![](`r my_top_tracks`){width=450px height=800px}

## Artist Frequency

Next, we have the `my_artist_frequency` function, which generates a bar graph of the user's most frequently listened to artists from the user's saved tracks or top tracks of all time. The function takes the user's category, vibe, name choice, and save location as arguments. It plots a bar graph with counts of songs from each of the top 10 artists and prints the name and number of the most frequent. The function will return the file path to the generated .png file.

Parameter options:

- `category`: "top" or "saved"

- `data`: the user's saved track data or top track data. If none specified, defaults to package data based on `category` arguments (`spotifywRapped::top_tracks_longterm` if `top` is specified).

- `vibe`: "soft", "neon", "neutral", or "bright"

- `name`: the name of the output file

- `saveto`: the directory to which the output file is saved


Here is an example of the artist frequency graphic generated for the user's top tracks:

```{r artist_frequency_top}

artist_frequency_top_demo <- my_artist_frequency(
  category = "top", vibe = "soft",
  name = "Top Tracks Artist Frequency",
  saveto = temp_save
)

```



![](`r artist_frequency_top_demo`){width=450px height=800px}


...and again for the user's saved tracks:

```{r artist_frequency_saved}
artist_frequency_saved_demo <- my_artist_frequency(
  category = "saved", vibe = "neon",
  name = "Saved Tracks Artist Frequency",
  saveto = temp_save
)
```


![](`r artist_frequency_saved_demo`){width=450px height=800px}

## Wordcloud - Frequented Artists

Finally, we have the `wordcloud` function, which generates a word cloud of the user's most frequently listened to artists from the user's saved or top tracks. The function takes the user's listening data, vibe, file name, and save location as arguments. It plots the word cloud over the backing graphic. The function will return the file path to the generated .png file. Note that the user must have phantom_js added to their PATH environment variable. This can easily be done using the `webshot` package with "webshot::install_phantomjs()".

Parameter options:

- `category`: "top" or "saved"

- `data`: the user's saved track data. If none specified, defaults to package data based on `category` arguments (`spotifywRapped::top_tracks_longterm` if `top` is specified. 

- `vibe`: "soft", "neon", "neutral", or "bright"

- `name`: the name of the output file

- `saveto`: the directory to save the output file

```{r wordcloud}
wordcloud_demo <- my_wordcloud(category = "long", vibe = "soft",
                               name = "wordcloud_top_artists",
                               saveto = temp_save)
```


![](`r wordcloud_demo`){width=450px height=800px}

## Behind the Scenes: Templates

Based on which functions are called and which inputs are given, the relevant Canva templates stored in the `inst/vibes` folder are called and embedded into a plot. 
For instance, a specification of `category = "songs"`,  `time = "long"`, and `vibe = "neon"` within the `my_top_five()` function would call the template titled `songs_long_neon.png`. Plot embedding is done using the `rasterImage` and `ggplot2` packages. Text is overlayed using the same packages, and the final plot is saved as a .png file using the .png graphics device.

## Datasets

The spotifywRapped package comes equipped with several example datasets that users can use to test the package's functions. These datasets were pulled from an author's personal Spotify account. The datasets are as follows:

- `saved_tracks`: a dataset containing the user's saved tracks from all time

- `top_tracks_longterm`: a dataset containing the user's top tracks from all time

- `top_tracks_mediumterm`: a dataset containing the user's top tracks from the past 6 months

- `top_tracks_shortterm`: a dataset containing the user's top tracks from the past 4 weeks

- `top_artists_longterm`: a dataset containing the user's top artists from all time

- `top_artists_mediumterm`: a dataset containing the user's top artists from the past 6 months

- `top_artists_shortterm`: a dataset containing the user's top artists from the past 4 weeks


## Limitations

- If users wish to use the `spotifywRapped` package for their own data, they must have a Spotify account and a Spotify Developer account. The package requires the user to have a Spotify API key to access their listening data. The package also requires the user to have preprocessed their data in the same way as outlined in the Data Processing section.

- Only 1000 observations were obtained for long-term top tracks and artists. Limits might be extended for medium and short term listening histories as well.

- Functionality is limited at this time to only 4 basic plotting options, but it could be expanded in the future. Note that to do so would require generating and adding additional background graphics to the `inst/vibes` folder within the internal package structure.

## Summary 

The spotifywRapped package provides a way for the user to visualize their listening history and patterns at any time (not just on a random day in December). The API data updates constantly, so users are able to get extremely up to date information when they generate their plots. The package uses numerical data and analysis within these graphics to further inform the user of their listening. They are also able to choose the vibe of these plots to enjoy unique and improved aesthetics compared to the basic Spotify Wrapped options. With energy vs. valence plots, top artist and track breakdowns, artist frequency plots, and frequent artist wordclouds, the possibilities are nearly endless. With spotifywRapped, Spotify Wrapped Day can be every day.
